### 2020-08-28 manuscript.tex
results update

Here, we present the \texttt{NarrowExchangeRate} (NER) operator. 
Let $r_A$, $r_B$, $r_C$, and $r_D$ be the clock rates of nodes $A$, $B$, $C$, and $D$, respectively. 
In addition to the modest topological change applied by \texttt{NarrowExchange}, NER also proposes new clock rates ${r_A}^\prime$, ${r_B}^\prime$, ${r_C}^\prime$, and ${r_D}^\prime$. While NER does not alter $t_D$ (i.e. ${t_D}^\prime \leftarrow t_D$), we also consider NERw - a special case of the NER operator which embarks $t_D$ on a random walk:

@@

\clearpage
\subsection*{A guided adaptive leaf rate operator}
\label{AVMVN_sect}

A \textit{guided} operator incorporates knowledge about neighbouring states, while an \textit{adaptive} operator undergoes a training process to improve its efficiency over time \cite{roberts2007coupling}. In previous work, parsimony scores and conditional clade probabilities of neighbouring trees have been employed by guided tree operators \cite{hohna2012guided,zhang2020using,meyer2019adaptive} and the latter has also been explored as the basis of  adaptive tree operators \cite{hohna2012guided,meyer2019adaptive}. The (adaptive) mirror kernel \cite{thawornwattana2018designing} learns a target distribution which acts as a `mirror image' of the current point $x$.  The adaptable variance multivariate normal (AVMVN) kernel \cite{baele2017adaptive,suchard2018bayesian} learns correlations between parameters during MCMC. Baele et al. 2017  observed a large increase ($\approx 5-10 \times$) in sampling efficiency from using the AVMVN kernel on clock rates and substitution model parameters across partitions \cite{baele2017adaptive}.

@@

This experiment revealed that nocons performed better than cons on datasets with small $L$ while cons performed better on larger datasets (\textbf{Fig. \ref{fig:round1Results}} and \textbf{\nameref{sect:SR_appendix}}). 
This result is unsurprising (\textbf{Fig. \ref{fig:rateparams}}).
Furthermore, the adapt setup dramatically improved mixing for \textit{real} by finding the right balance between cons and nocons.
This yielded an average ESS/hr of XXX times that of cons and YYY times that of nocons, with respect to leaf branch rates.   
However adapt did not improve \textit{cat} or \textit{quant}, suggesting that the default operator weighting scheme was sufficient in either case.
Overall, the \texttt{AdaptiveOperatorSampler} operator will be included in all subsequent rounds in the tournament.

@@

\clearpage
\subsection*{Round 2: The \textit{real} parameterisation yields the fastest mixing }

@@

We benchmarked the adapt (\textit{real}) configuration with a) standard uniform proposal kernels, and b) Bactrian(0.95) kernels.
These kernels applied to all clock rate related operators (\textbf{Table \ref{table:bactriankernels}}).

@@

>>>>

Here, we present the \texttt{NarrowExchangeRate} (NER) operator. 
Let $r_A$, $r_B$, $r_C$, and $r_D$ be the substitution rates of nodes $A$, $B$, $C$, and $D$, respectively. 
In addition to the modest topological change applied by \texttt{NarrowExchange}, NER also proposes new branch rates ${r_A}^\prime$, ${r_B}^\prime$, ${r_C}^\prime$, and ${r_D}^\prime$. While NER does not alter $t_D$ (i.e. ${t_D}^\prime \leftarrow t_D$), we also consider NERw - a special case of the NER operator which embarks $t_D$ on a random walk:

@@

\clearpage
\subsection*{An adaptive leaf rate operator}
\label{AVMVN_sect}

%, such as \texttt{AdaptiveOperatorSampler},
Adaptive operators undergo a training process to improve their efficiency over time \cite{roberts2007coupling}. 
For example, the conditional clade probabilities of neighbouring trees can serve as the basis of adaptive tree operators \cite{hohna2012guided,zhang2020using,meyer2019adaptive},
and the mirror kernel learns a target distribution which acts as a `mirror image' of the current point $x$ \cite{thawornwattana2018designing}.  
The \texttt{AdaptiveOperatorSampler} operator, introduced in this article, learns the weight of sub-operators.
Finally, the adaptable variance multivariate normal (AVMVN) kernel  learns correlations between parameters during MCMC \cite{baele2017adaptive,suchard2018bayesian}. Baele et al. 2017  observed a large increase ($\approx 5-10 \times$) in sampling efficiency from using the AVMVN kernel on clock rates XXX(check)XXX and substitution model parameters across partitions \cite{baele2017adaptive}.

@@

This experiment revealed that nocons performed better than cons on datasets with small $L$ while cons performed better on larger datasets (\textbf{Fig. \ref{fig:round1Results}} and \textbf{\nameref{sect:SR_appendix}}). 
This result is unsurprising (\textbf{Fig. \ref{fig:rateparams}}).
Furthermore, the adapt setup dramatically improved mixing for \textit{real} by finding the right balance between cons and nocons.
This yielded an ESS/hr (averaged across all 9 datasets) 80\% faster than cons and 120\% faster than nocons, with respect to leaf branch rates, and 250\% and 1800\% faster for $\sigma$.
Similar results were observed with \textit{quant}.
However adapt neither helped nor harmed \textit{cat}, suggesting that the default operator weighting scheme was sufficient.
Overall, the \texttt{AdaptiveOperatorSampler} operator will be included in all subsequent rounds in the tournament.

@@

\clearpage
\subsection*{Round 2: The \textit{real} parameterisation yielded the fastest mixing }

@@

We benchmarked the adapt (\textit{real}) configuration with a) standard uniform proposal kernels, and b) Bactrian(0.95) kernels.
These kernels applied to all clock model related operators (\textbf{Table \ref{table:bactriankernels}}).

@@

Our experiments confirmed that $\text{NER}\{\mathcal{D}_{AE}, \mathcal{D}_{BE}, \mathcal{D}_{CE}\}$ is indeed superior on larger datasets (where $L > 1$kb; \textbf{Fig. \ref{fig:round4Results}}).
Whereas, on smaller datasets the standard \texttt{NarrowExchange} operator had a higher acceptance rate.
Using an adaptive operator (\texttt{AdaptiveOperatorSampler}) removes the burden from the user in selecting which operator is the best pick.
The average acceptance rate of

\begin{figure}[!h]
\includegraphics[width=\textwidth]{benchmarking/benchmarkingVM/ESS_round4.pdf}
\caption{\textbf{Round 4: benchmarking the NER operators.} 
The learned weights (left) behind the two NER operators ($\text{NER} \{\}$ and $\text{NER}\{\mathcal{D}_{AE}, \mathcal{D}_{BE}, \mathcal{D}_{CE}\}$), and the relative difference between their acceptance rates $\alpha$ (right), are presented as functions of sequence length.
Logistic and linear regression models are shown, respectively. }
\label{fig:round4Results}
\end{figure}
